BEM150 Final Project
========================================================
Eli Alster  
Vansh Kumar  
Angad Rekhi  

     We analyzed the NYC traffic data set. In order to simplify analysis, we only considered the 2013 data so the data set would be fixed.
  
```{r LoadData, echo=FALSE, cache=FALSE}
     library("plyr")
     library("hexbin")
     #install.packages('e1071', dependencies = TRUE)
     library(e1071)
     
     # Read raw data
     df_main = read.csv("collisions.csv", header = TRUE, sep="\t")

     # Extract training and test sets and replace NA's with 0
     df_main[is.na(df_main)] = 0
     df_used = subset(df_main, (year == 2012 | year == 2013))

     df_used$lat = as.numeric(as.character(df_used$lat))
```

```{r accident_plot, error=TRUE, cache=TRUE}
     df_in = subset(df_used, lat < 44)
     bin = hexbin(df_in$lon, df_in$lat, xbi=75)
     plot(bin, main="2013 Traffic Deaths in NY by Location", xlab="Longitude", ylab="")
     mtext(text = "Latitude", side=2, line=3)
```
Accident frequency map!

Histogram of reason for accident vs total people killed.

```{r Reasons_of_Death, cache=TRUE, error=TRUE}
     reasons = df_used[,39:68]

     mat = matrix(nrow=4, ncol=ncol(reasons))
     for ( r in 1:ncol(reasons) ) 
     {
          reason = names(reasons)[[r]]
          reason_row = subset(df_used, df_used[[reason]]==TRUE)
          ped = sum(reason_row[,"pedestr_killed"])
          pass = sum(reason_row[,"passengers_killed"])
          motor = sum(reason_row[,"motorists_killed"])
          cycle = sum(reason_row[,"cyclists_killed"])
          
          mat[,r] = c(ped, pass, motor, cycle)
     }

     par(mar=c(4,10,2,1)) # increase margins on plot 
     par(las=2)          # align accident names horizontally
     colors=c("blue","green","red","black")
     barplot(mat, horiz=TRUE, names.arg=colnames(reasons), col=colors, main="2013 NYC Deaths by Kind of Accident", xlab="Number of Deaths", xlim=c(0,40), cex.names=0.5)
     legend(x=25,y=25,fill=colors,legend=c("Pedestrians","Passengers","Motorists","Cyclists"))
 
```

Death Predictions
------------------

In this project, we decided to see if we could predict if people would die in an accident. This would be useful for emergency response dispatchers. Given the location of the accident and the situation in which the accident takes place, we used an svm classifier to determine whether deaths would occur. Dispatchers could use this prediction algorithm to prioritize resources.

``` {r Data_Munging, eval = TRUE, cache=TRUE}

# Notes:
# x = classification matrix
# y = factors for each type
# scale = FALSE if just factors
# 39:68 reason, lon, lat, 22:38 vehicle type

# Separate training and test sets
df_training = subset(df_used, year == 2012)
df_test = subset(df_used, year == 2013)

# Make training set more balanced
# About 3/4 of the accidents had no injuries; so for each accident, delete it with probability 2/3 to roughly balance the training set
random = runif(nrow(df_training), 0, 1)
df_training = df_training[!(df_training$total_injured==0 & random<2/3),]

# Training Set
#    Extract explanatory variables
training_x = data.frame(df_training$lon, df_training$lat, df_training[,39:68], df_training[,22:38])
#    Extract dependent variable
training_y = as.numeric(df_training$total_injured != 0)  # note: vector, not data frame

# Make training set more balanced 

# Test Set
#    Extract explanatory variables
test_x = data.frame(df_test$lon, df_test$lat, df_test[,39:68], df_test[,22:38])
test_x[is.na(test_x)] = 0
#    Extract dependent variable
test_y = as.numeric(df_test$total_injured != 0)  # note: vector, not data frame
test_y[is.na(test_y)] = 0

# Train SVM
model = svm(x = training_x, y = as.factor(training_y), kernel = "linear", cost = .1, type = "C-classification", scale = FALSE)

# Test SVM
predictions = predict(model, test_x, decision.values = TRUE)
predictions_2 = as.numeric(as.character(predictions))

# See how many were correct
trials = nrow(test_x)
correct_amt = sum(test_y == predictions)
correct_pct = correct_amt / trials
print(correct_pct)

```

The classifier is `r correct_pct*100` % accurate. We used a training set of 2012 data and a test set of 2013 accident data.

```{r linear_regression, error=TRUE}

# Training Set
# 
# training_x is a data.frame
# Includes: df_training$lon, df_training$lat, df_training[,39:68], df_training[,22:38]
# Those other columns are: type of vehicle and reason listed for accident
# training_y numeric vector is boolean indicator for injury
# 
# Similarly for test_x, test_y

# Simple Linear Model
training_xm = as.matrix(training_x)
#training_xm1 = cbind(rep(x = 1, times = nrow(training_x)), training_xm) #add 1 column
model_linear = lm(training_y ~ training_xm)
# NOTE!!!: This gives 9 problematic variables which give coefficients of NA.
model_linear$coefficients[is.na(model_linear$coefficients)] = 0

# To get threshold value, find midpoint of min and max of training-x
training_xm1 = cbind(rep(x = 1, times = nrow(training_xm)), training_xm)
predictions_training = training_xm1 %*% model_linear$coefficients
thresh = mean(c(min(predictions_training), max(predictions_training)))

# Test Linear
test_x_mat = as.matrix(test_x)
test_x_mat = cbind(rep(x = 1, times = nrow(test_x_mat)), test_x_mat) # add column of ones
predictions_linear = test_x_mat %*% model_linear$coefficients

# Convert regression results to actual predictions
predictions = as.numeric(predictions_linear > thresh)

# See how many were correct
trials = length(predictions)
correct_amt = sum(test_y == predictions)
correct_pct = correct_amt / trials
print(correct_pct)
```

```{r Neural Net}

#install.packages("neuralnet", dependencies=TRUE)
library("neuralnet")

# Training Set
#    Extract explanatory variables
NN_training_x = training_x
#    Extract dependent variable
NN_training_y = data.frame(training_y)# note: vector, not data frame
colnames(NN_training_y) <- "y"
train = cbind(NN_training_y, NN_training_x)

# Test Set
#    Extract explanatory variables
NN_test_x = test_x
#    Extract dependent variable
NN_test_y = data.frame(test_y)
test = cbind(NN_test_y, NN_test_x)

# Create formula that describes our neural net and train it
n <- names(train)
f <- as.formula(paste("y ~", paste(n[!n %in% "y"], collapse = " + ")))
NN = neuralnet(f, data = train)

# Compute results of the neural network on our test data
results <- compute(NN, data.frame(test))
ls(results)
print(results$net.result)

```